---
title: "Exploration of the use of the National Vegetation Database to estimate vegetation parameters for hydrological modeling"
author: "Jasper Slingsby"
date: "8 Novemer 2016"
output: html_document
---

```{r, echo=FALSE}
#suppressWarnings(suppressMessages(
suppressWarnings(suppressMessages(library(gdata, quietly=T)))
suppressWarnings(suppressMessages(library(sp, quietly=T)))
suppressWarnings(suppressMessages(library(maps, quietly=T)))
suppressWarnings(suppressMessages(library(maptools, quietly=T)))
suppressWarnings(suppressMessages(library(raster, quietly=T)))
suppressWarnings(suppressMessages(library(ggplot2, quietly=T)))
suppressWarnings(suppressMessages(library(gstat, quietly=T)))
suppressWarnings(suppressMessages(library(rgdal, quietly=T)))
suppressWarnings(suppressMessages(library(gjam, quietly=T)))
suppressWarnings(suppressMessages(library(ggmap, quietly=T)))
suppressWarnings(suppressMessages(library(dplyr, quietly=T)))
#))

###Set data working directory
datwd <- "/Users/jasper/Documents/Databases/NVD/NVD_16_09_2016/"

##########################################
###2) Get data
##########################################

#Get vegmap shapefile
#veg06 <- readOGR(dsn = "/Users/jasper/Documents/GIS/VegToolsRaw/VM2006/vegmap2006_geo.shp", layer = "vegmap2006_geo", verbose = FALSE) #Vegmap 2006
suppressMessages(veg <- readOGR(dsn = "/Users/jasper/Documents/GIS/VegToolsRaw/VM2012/nvm2012beta2_wgs84_Geo.shp", layer = "nvm2012beta2_wgs84_Geo", verbose = FALSE)) #Vegmap 2012

#Get hydrology veg clusters and merge into vegmap data
hydC <- read.xls("/Users/jasper/Documents/GIS/VegToolsRaw/Hydrology_Baseline/Final clusters.xls", sheet=1) #Hydrology clusters from Michele Toucher 1 Nov 2016
names(hydC)[which(names(hydC)=="Name")] <- "NAME"
hydC <- hydC[,c(1,8)]

x <- merge(veg@data, hydC, all.x=T, sort=F)
x <- x[order(x$OBJECTID_1),]
veg@data <- x

#Get pentads
#suppressWarnings(pen <- readOGR(dsn = "/Users/jasper/Documents/GIS/VegToolsRaw/pentads/pentads/PENT_PRO.SHP", layer = "PENT_PRO"))

#Get landcover (lumped to natural vs not natural classes)
lc <- raster("/Users/jasper/Documents/GIS/VegToolsRaw/Landcover/LCsmiso_trans1.tif")

#Get rePhoto database
#rP <- read.csv("/Users/jasper/Dropbox/SAEON/Grants/FBIP2016/rePhoto/Index to database 16.iv.2013.csv", stringsAsFactors = F) 
#Get NVD
#HD <- read.csv(paste0(datwd, "HeaderData_NVD2014.txt")) #Header Data
HD <- read.csv(paste0(datwd, "HeaderData_NVD2014.csv")) #Header Data
#HD2 <- read.xls(paste0(datwd, "HeaderData_NVD2014.xlsx"), sheet=1) #Header Data
SD <- read.csv(paste0(datwd, "SpeciesData.txt")) #Species Data
SN <- read.csv(paste0(datwd, "SpeciesList.txt")) #Species Names
MD <- read.csv(paste0(datwd, "MetadataActionsEtc_Head.txt")) #Meta Data

#Get SANBI checklist
#clist <- read.xls("/Users/jasper/Documents/GIS/VegToolsRaw/SANBI_checklist_20June2016.xlsx", sheet = 1)

##Get POSA
#posa <- read.csv("/Users/jasper/Documents/Databases/POSAoffline/POSAoffline_Species_GrowthForm.txt")
#posaN <- read.csv("/Users/jasper/Documents/Databases/POSAoffline/POSAoffline_Species.txt")

#posa <- merge(posaN, posa, all.x=T)
#posa <- posa[-which(posa$GrowthForm %in% c("Bryophyte", "Lichen")),]

#posa$NameCode <- paste(posa$GENNO, posa$SPNO, sep = "_")
#gfs <- as.data.frame(table(posa$NameCode, posa$GrowthForm))

#write.csv(levels(posa$GrowthForm), "posa_growth_forms.csv")

#colnames(x) <- c("plot", "abund", "id")
#y <- tapply(x$abund, list(x$plot, x$id), sum)
#y[is.na(y)] <- 0
#as.data.frame(y)

##########################################
###3) Summary of NVD fields
##########################################

HeadSummary <- data.frame(
  Complete = sapply(HD, function(x){sum(!is.na(x))}), #Non-NA values
  LengthUniqueVals = sapply(HD, function(x){length(levels(as.factor(x)))}), #Number of unique values or factor levels
  UniqueVals = sapply(HD, function(x){if(length(levels(as.factor(x)))<100) {paste(levels(as.factor(x)), collapse=",")} else {">100"}}))
#write.csv(HeadSummary, paste0(datwd,"SummaryStats.csv"))

##########################################
###4) Crosscheck species names with checklist
##########################################

#clist$CHECKLISTNAME <- paste(clist$genus, clist$sp1, sep = " ")
#snr <- merge(SN, clist, by.x = "SPNUMBER", by.y = "spnumber", all.x = T)

#snr <- snr[,c("SPNUMBER", "SPECIES_NR", "SPECIESNAME", "CHECKLISTNAME", "syncat", "synof")]

#SnD <- merge(SD, snr)

#SnD <- merge(SnD, posa)

#snr <- SN[,c("SPECIES_NR", "SPNUMBER")]
#snr <- merge(SD, snr)
#sum(snr$SPNUMBER %in% clist$spnumber)

##########################################
###4) Visualize plots and repeat photos
##########################################

#p <- rP
#suppressWarnings(suppressMessages(p$DD..S. <- -as.numeric(p$DD..S.)))
#suppressWarnings(suppressMessages(p$DD..E. <- as.numeric(p$DD..E.)))
#p <- p[!is.na(p$DD..S.) & !is.na(p$DD..E.),]
#coordinates(p) <- cbind(p$DD..E., p$DD..S.)
#proj4string(p) <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0" #Assign a coordinate reference system (Geographic WGS84)

#Make new data frame to use for spatial analyses
d <- HD[!is.na(HD$Lat) & !is.na(HD$Long),] # remove plots with no Lat or Long

#Convert data frame to a SpatialPointsDataFrame
coordinates(d) <- cbind(d$Long, d$Lat)
proj4string(d) <- "+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0" #Assign a coordinate reference system (Geographic WGS84)

#Fix errors by overlaying points and polygon
map <- map("world", region = "south africa", fill = TRUE, plot=F)
IDs <- sapply(strsplit(map$names, ":"), function(x) x[1])
map <- map2SpatialPolygons(map, IDs=IDs, proj4string=CRS("+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0"))
map <- crop(map, d) #Crop the map to the points (to drop the Prince Edward Islands)

#Trim plots with poor locality data
#d <- d[-which(d$Loc_Conf<0),]
#d <- d[-which(is.na(d$Loc_Conf)),]

#Extract vegmap, hydrological clusters and land cover data
x <- d %over% map
v <- d %over% veg
d$Veg <- v$NAME
d$Hyd <- v$Cluster.name
suppressMessages(d$LandCover <- extract(lc, d)) #1 = Natural, 2 = Transformed

#d <- d[which(x==1), ]
#x <- d %over% map
#y <- p %over% map

#Insert frequency of plots into vegmap
vpf <- table(d@data$Veg)
vpf <- as.data.frame(vpf)
names(vpf)[1] <- "NAME"

vpf <- merge(veg@data, vpf)
vpf <- vpf[order(vpf$OBJECTID_1),]
veg@data <- vpf

```

**Lets have a look at the basic stats**

Number of plots altogether = `r  nrow(HD)`

Number of plots with spatial data = `r nrow(d)`

Number of unique species names = `r length(unique(SN$LETTERCODE))` - lots if issues there!

Number of unique genus names = `r length(unique(SN$GenName))` - suspicious too...

Number of species by site observations = `r nrow(SD)`

***

Ok, lets look at the spatial distribution with sites with larger circles indicating poor location confidence, and red sites falling in the sea:

```{r, echo=FALSE}
plot(map)
suppressWarnings(points(d[which(x==1), ], cex=.1*log(d[which(x==1), ]$Loc_Conf+1))) #Plots
points(d[is.na(x),], cex=.25, col="red") #Plots at sea (or next door)
```

***

What about representation of plots in the National Vegetation Map (2012) vegetation types (447 altogether)?

```{r, echo=FALSE}
hist(table(d$Veg), ylab = "Number of vegetation types", xlab = "Number of plots", main ="")
```

There are `r sum(table(d$Veg)<1)` veg types without any plot data, `r sum(table(d$Veg)<5)` veg types with less than 5 plots, `r sum(table(d$Veg)<10)` veg types with less than 10 plots...

***

What about representation of plots in hydrological clusters (veg types lumped into 132 clusters based on similar climates)?

```{r, echo=FALSE}
hist(table(d$Hyd), ylab = "Number of hydro clusters", xlab = "Number of plots", main ="")
```

There are `r sum(table(d$Hyd)<1)` veg types without any plot data, `r sum(table(d$Hyd)<5)` veg types with less than 5 plots, `r sum(table(d$Hyd)<10)` veg types with less than 10 plots...

***

But! `r summary(as.factor(d$LandCover))[2]` of the `r nrow(d@data)` plots with spatial information are now in transformed landscapes according to the 2013 National Land Cover. So let's look at all those stats again, excluding transformed plots...

```{r, echo=FALSE}
nd <- d[which(d$LandCover==1),]
hist(table(nd$Hyd), ylab = "Number of hydro clusters", xlab = "Number of plots", main ="")
```

There are `r sum(table(nd$Hyd)<1)` veg types without any plot data, `r sum(table(nd$Hyd)<5)` veg types with less than 5 plots, `r sum(table(nd$Hyd)<10)` veg types with less than 10 plots...

We can still use the data from the transformed sites, but we obviously can't revisit these sites to measure any additional parameters.

***
Here's teh representation by ecohydrological cluster:

```{r, echo=F}
table(d$Hyd)

# Plot VegMap
#plot(veg, col=rev(heat.colors(veg@data$Freq/veg@data$SHAPE_AREA)), border="transparent")
#points(nd)


#Veg map data for plotting in ggplot2
#vd <- tbl_df(as.data.frame(veg))
#vvd <- fortify(veg)
#vd$id <- as.character(0:(nrow(vd)-1))
#vvd <- left_join(vvd, vd, by="id", copy=T)
#m <- ggplot() + geom_polygon(data = vvd, aes(x = long, y = lat, group = id, color = Freq), size = 0.25) + coord_map()
#m
```

***